---
title: "TraVisPies_notebook"
author: "Sam De Craemer"
date: "`r Sys.Date()`"
# output: html_TraVisPies_notebook
---

# Description

This code aims to produce the pie charts plots proposed in the article TraVis Pies: A Guide for Stable Isotope Metabolomics Interpretation Using an Intuitive Visualization, based on 3 input types: metadata, raw abundance and fractional contribution. All of this can also be done for data including more than one tracer, and up to two cohort factors

The plots are generated using ggplot2 for each compound in the input data several other tidyverse packages are used to get the data in the right format to generate the required figures.

## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load functions and libraries

```{r functions and libs, message=FALSE}
# Functions and libraries ---------------------------------------------------------------
#libraries for functions used
library(here)    #to make r source from file location instead of magic stuff
library(vroom)   #for error messages on box
library(readxl)       #for reading excel files
library(forcats)      #for factor manipulation
library(dplyr)        #for faster.easier manipulation of data
library(tibble)       #for manipulating tibbles
library(readr)        #for writing .csv file of merged output
library(tidyr)        #for restructuring data tibbles
library(broom)        #for using regression models in dplyr pipes
library(ggplot2)      #for generating the pie chart plots

#load functions to support the app 
source(here::here("Functions and modules/metabolomics_tibble_functions.R"))
source(here::here("Functions and modules/TraVis_Pies_functions.R"))
```

# Load data and put settings

```{r input general and filepath, message=FALSE}
#Data file locations and specifications
#rawpath is windows copied folder with files, need this command to properly 
#read in R without having to modify the strings by hand
#example: rawpath<-r"(C:\User\Projects\Pie charts)"
#can also specify relative to the project folder using the here::here command

#set optional variables to NULL, functions designed to handle absence
isostring<-comparative_factor_column<-sampletype_column<-factor_column <- norm_column <- 
  tracer_column<-inputtype<-libfile<-lib_tb<-factor_levels_ordered<-col_labeling<-NULL

#set required variables to a default value that can be changed for specific 
#projects further in input
metastring<-"meta"
abundstring<-"abund"
labelstring<-"iso"
minfract_detected<-0

#set filepath and check metadata
path<-here::here("Example_data/Experimental examples for nonUI app/Excel_1factor_1tracer")
savepath<-path
excelfile<-"example excel.xlsx"
inputpath<-paste(path,excelfile,sep = "/")
read_excel(inputpath,
           which(grepl("meta",tolower(excel_sheets(inputpath)))))
```

```{r input metadata}
#set metadata variables
sample_column <-"Sample"
factor_column <- "cohort"   #"None" if not present, or 1 or two element vector
norm_column <- "Normalisation"   #"None" if not present
sampletype_column<-"sample_type"
(factor_levels<-read_excel(inputpath,
           which(grepl("meta",tolower(excel_sheets(inputpath)))))%>%
  pull(factor_column)%>%
  unique())

#set factor levels variables
factor_levels_ordered<-factor_levels[2:4]

#make vector of factors
factor_columns <- c(factor_column,comparative_factor_column)

print(paste0("Selected levels in order as given in factor_levels_ordered: ",paste0(factor_levels_ordered,collapse = ", ")))
```

```{r labeling check}
#set 
isostring<-"_C13-0"
read_excel(inputpath,
           which(grepl("iso",tolower(excel_sheets(inputpath)))))%>%
  colnames()
```

```{r libfile, message=FALSE, warning=FALSE}
#load file or quote if not required so it remains NULL
libfile<-"Lib_excelexample.csv"
lib_tb<-vroom::vroom(paste0(path,"/",libfile),delim = ",")
```

```{r input piecharts}
#Miscellaneous
P_isotopologues<-T                    #leave at false, only input fraction contribution data 
log_abund<-F
detail_charts<-T                      #makes images with detail for solo use
pathway_charts<-T                        #makes images fit for pathway 
iso_charts<-F
normalize<-(length(norm_column)>0)    #normalize abundances?
print_tables<-F                       #print generated tables to console?
compounds<-NULL                       #which compounds included; NULL => all
show_P<-T                             #show P values on pie plots

#figure appearance parameters if not specified in project
col_labeling<-c("#ffd966","#bfbfbf")   #colors for labeled and unlabeled fractions, NULL to use default color scheme 
maxcol_facet<-3                       #maximum amount of images horizontal
include_name<-T                        #include compound name on figure
include_legend<-T                      #include legend on figure

#axis names and fonts, load font library
#Font: set to "sans" to use standard font. For other available options run
#windowsFonts() after loading the extrafont library.If desired font not present,
#check link below on importing fonts:
#https://www.r-bloggers.com/2013/02/change-fonts-in-ggplot2-and-create-xkcd-style-graphs/
xAxLab<-""                            #xlabel, best "" for grids of pies
yAxLab<-""                            #ylabel, best "" for grids of pies
font<-"sans"

#figure size parameters in cm for detailed pie charts. A4 landscape is 
#recommended: width=24.6 and height 16
width<- 24.6                          
height<-16                            

#fontsizes on  detailed pie charts: Cohort names above chart, legend names above legend, all others separate)
#when width 24.6=height=16 and font= calibri
# for 2 cohorts 28 24 24 recommended
# for 3 cohorts 18 16 16 recommended
cohortsize<-18                        #text size of cohort names
legendtitlesize<-16                   #set legend font size
otherfontsize<-12                     #adapt text size of all but those above


#fontsizes on summary pie charts for pathway
#when width 6.15=height=2.81 and font= calibri
# for 2 cohorts 28 24 recommended
# for 3 cohorts 24 20 recommended
mapcohortsize<-16
mapotherfontsize<-18

#textlabel fractional contribution parameters
#FC_position sets where FC should be displayed. "center" to display in center,
#"slice" to display in labeled slice. If more than one tracer is used, always
#defaults to slice.
#min_lab_dist sets minimal distance at which FC label is plotted. 0 is center
#1 is the outer circle. If distance would be smaller based on pie abundance,
#label is plotted at min_lab_dist  distance from the circle center
FC_position<-"center"  
min_lab_dist<-0.7                       
label_decimals<-1                     #amount of decimals in FC label
percent_add<-T                         #if true adds "%" to the FC label

#linetypes and colour of concentric circles
# circlelinetypes<-c(3,4,2,6)  #concentric lines from inside to out dotted, dot dash, dash, doubledash
# circlelinetypes<-c(3,3,3,3)  #all concentric lines dotted
# circlelinetypes<-c(2,2,2,2)  #all concentric lines dashed
circlelinetypes<-c(1,1,1,1)  #all concentric lines solid
# circlelinetypes<-c(0,0,0,0)  #no concentric circles

circlelinecolor<-"gray"

#Other settings
alpha<-0.7
format<-"png"
```

# Make standardized merged tibble

```{r make merged tibble,echo=FALSE}
tb<-dolly_to_longtibble(path,inputpath,metastring = metastring,
                        abundstring = abundstring,labelstring = labelstring,
                        isostring = isostring,
                        sample_column = sample_column,
                        factor_column = factor_column,
                        comparative_factor_column =
                          comparative_factor_column,
                        factor_levels_ordered = factor_levels_ordered,
                        norm_column = norm_column,
                        tracer_column = tracer_column,
                        sampletype_column = sampletype_column,lib_tb = lib_tb,
                        savedata=T)

tb
```

# Check FC plot

```{r check FC pie, echo=FALSE}
checkplot<-plot_fromtibble(tb,charttype="FC",compound_col="compound",
                selected_compound=unique(tb$compound)[6],
                factor_columns=factor_columns,
                tracer_column=tracer_column,
                log_abund=log_abund,
                circlelinecolor=circlelinecolor,
                circlelinetypes=circlelinetypes,
                maxcol_facet=maxcol_facet,
                include_name=include_name,
                col_labeling=col_labeling,
                alpha=alpha,
                otherfontsize=otherfontsize,
                font=font,
                legendtitlesize=legendtitlesize,
                cohortsize=cohortsize,
                include_legend=include_legend,
                show_P=show_P)
checkplot
```

# Check iso plot if desired

```{r check iso pie, echo=FALSE}
if (iso_charts){
  checkplot<-plot_fromtibble(tb,charttype="iso",compound_col="compound",
                selected_compound=unique(tb$compound)[6],
                factor_columns=factor_columns,
                tracer_column=tracer_column,
                log_abund=log_abund,
                circlelinecolor=circlelinecolor,
                circlelinetypes=circlelinetypes,
                maxcol_facet=maxcol_facet,
                include_name=include_name,
                col_labeling=col_labeling,
                alpha=alpha,
                otherfontsize=otherfontsize,
                font=font,
                legendtitlesize=legendtitlesize,
                cohortsize=cohortsize,
                include_legend=include_legend,
                show_P=show_P)
checkplot
}

```

# Save plots with settings above

```{r generate and save pies}
mergedtibble_to_pies(tb,compound_col="compound",
                     detail_charts=detail_charts,
                     pathway_charts=pathway_charts,
                     iso_charts=iso_charts,
                     savepath=savepath,
                     normalize=normalize,
                     factor_columns=factor_columns,
                     tracer_column=tracer_column,
                     log_abund=log_abund,
                     circlelinecolor=circlelinecolor,
                     circlelinetypes=circlelinetypes,
                     maxcol_facet=maxcol_facet,
                     include_name=include_name,
                     col_labeling=col_labeling,
                     alpha=alpha,
                     otherfontsize=otherfontsize,
                     font=font,
                     legendtitlesize=legendtitlesize,
                     cohortsize=cohortsize,
                     include_legend=include_legend,
                     format=format,
                     show_P=show_P,
                     width=width,height=height)
```

## 
